---
title: "Hello R"
author: "Andre Sacilotti"
date: "21/05/2020"
output: html_document
---

```{r install, include=FALSE}
library(knitr)
library(rmarkdown)
library(dplyr)
library(tidyverse)

setwd('C:/Users/Andre/R/Hello Word')
```

# Basic Algebra

Learning some basic operations

```{r}
2+3
18/25
14/6+3
14/(6+3)
25+36/2
(25+36)/2
2^2
4^(1/2)
27^(1/3)
```

# Number Notation

In R are used the notation $ey$, that is the same as $10^y$, for numbers with more than 5 digits, as following

```{r}
10000
100000
0.001
0.0001
```

# Variables

Like in programming languages like python, or lua, we don't need to declare the type of the variable

```{r}

X = 10
Y = 20

Z = X*Y

Z

name = "Andre"
aux = "Hello"

paste(aux, name, sep = " ")
```

# Vectors

R has some basic vectors created with c(), like following

```{r}
vec = c(1,2,3,4,5,6,7,8)
vec
```

```{r}
vec2 = c(.0005, .540, .324, .1, .0001556)
vec2
```

```{r}
vec3 = c(vec, vec2)
print(vec3)
```

But we can create Sequences, using X:Y, quere X is the initial value and Y the last value from the sequence, look:

```{r}
3:27

```

Or yet using the seq() command

```{r}
seq(from = 2, to = 50, by = 3)

```


Or even with rep() command

```{r}
rep(x = c(1,2,3), times = 2)
```

```{r}
rep(x = c(1,2,3), each = 2)
```

```{r}
rep(x=c(1,2,3), times=2, each=2)
```

We can sort the vector

```{r}
sort(rep(x=c(1,2,3), times=2, each=2), decreasing = FALSE)
```
```{r}
sort(rep(x=c(1,2,3), times=2, each=2), decreasing = TRUE)
```

Measure the lenght of the vector

```{r}
length(sort(rep(x=c(1,2,3), times=2, each=2), decreasing = FALSE))
```

R had something different from others programming language, here the vector index starts from 1 (WHY???? WHY???????)

```{r}
vec[1]
```

And we can pass an vector with index that we want, repeting the index or even and out of index number, that will return NA

```{r}
vec[c(1,1,9,4)]
```

And someone that programs in python always uses things like list[-1], but, in R its delete the first element, look:

```{r}
vec
```
```{r}
vec[-1]
```
```{r}
vec
```

# Matrix

Giving the number of rows and columns, we can build an matrix

```{r}
matrix(data = c(1,2,3,4), ncol = 2, nrow = 2)
```

If we have 2 vectors or more, R has a function to build an matrix using the vectors as columns or rows, look:

```{r}
rbind(1:3,4:6)
```

```{r}
cbind(1:3,4:6)
```

The dim, nrow, ncol commands can return the required dimension of the matrix look
```{r}
aux = rbind(1:3,4:6)
print(dim(aux))
print(nrow(aux))
print(ncol(aux))
```

Like vectors, using [] we can return some elements of the matrix

```{r}
aux[2,]
```

```{r}
aux[,2]
```

```{r}
aux[2]
```

Transpose Matrix

```{r}
t(aux)
```

Identity Matrix

```{r}
A = diag(x=3)
A
```

Dot product

```{r}
2*aux
```

Matrix multiplication

```{r}
A = matrix(data = c(1,2,3,4), ncol=2, nrow=2)
B = matrix(data = c(4,5,6,7), ncol=2, nrow=2)

print(A)
print(B)
```

```{r}
A%*%B
```

```{r}
B%*%A
```

Inverse Matrix

```{r}
solve(A)
```

```{r}
solve(A)%*%A
```

Multidimensional Matrix, using the array command the dim argument are (nrow, ncol, dimension1, ...., dimensionN)

```{r}
BR = array(data=rep(1:24,times=3),dim=c(3,4,2))
BR
```

```{r}
BR = array(data=rep(1:24,times=3),dim=c(3,4,2,2,2))
BR
```

Which command help search with a condition inside matrix, vector, array...

```{r}
x = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20)
which(x = x>5)
```
We can find using which as index

```{r}
vec = c(1,2,3,4,5,6,7,8)
print(vec)
vec[which(x=vec<3)]
```

And we can delete, using -which
```{r}
vec = c(1,2,3,4,5,6,7,8)
print(vec)
vec[-which(x=vec<5)]
```



# Conditional Operator

```{r}
(TRUE&&FALSE)||(TRUE&&FALSE)
```

```{r}
(6< 10) && (6< 8)
```

# DataFrame

Yay! This is really nice way to present and use data, similar to python-pandas dataframe

```{r}
df = data.frame(Name = c("Carl", "Roberto", "Andre"), Age = c(25, 32, 18), Country = c("USA", "CAN", "BR"),
                Sex = c("F", "M", 'M'), Sport = c("Golf", "Swiming", "Run"))
df
```

```{r}
df[1]
```

Using df[X , c()] we can get the coluns c() of the row X, or if we use df[,c()] we will get all the rows with only c() columns 

```{r}
df[1,c(1,4)]
```

Using df[X, Y], X means the row and Y the columns

```{r}
df[2,1]
```

With the name of the columns, using X$Y, it returns the all values from the column Y

```{r}
df$Age
```

Of course, we can add dataframes

```{r}
new = data.frame(Name="Brian",Age=7,Sex="M", Country="USA", Sport="NA" )
new
```

```{r}
df = rbind(df,new)
df
```

And finally, filter elements in dataframe giving some condition

```{r}
df[df$Age >= 18,]
```


```{r}
df[df$Country =="USA",]
```

# Loops

R loops are simple like in python but uses brackets like C++

```{r}
for(i in 1:10){
  print(i)
}
```

```{r}
i=0
while (i < 5){
print(i)
i <- i + 1
}
```


# If condition

```{r}
x = 25
if(x < 10){
  print('less')
} else{
  print('greater')
}
```

# Functions

```{r}
Sum <- function(x,y){
  return(x+y)
}

Sum(1,3)
```

# Loading CSV or Datasets

Built-in dataset

```{r}
library(datasets)
data(iris)
summary(iris)

```



