---
title: "Hello R"
author: "Andre Sacilotti"
date: "21/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Iris Data Analysis

This is my first project in R, i started my studies recently, so i will apply everything i have learned so far

### Setup

```{r}
#install.packages('e1071')
#install.packages("caret")
library(datasets)
library(e1071)
library(caret)
```

### Reading the Dataset

Look to the dataframe of an subset of the iris data

```{r}
df = iris
head(df)
```

## Exploratory Data Analysis

Look bellow, using dim(), it returned that the dataframe has 150 elements and 5 columns. And using the summary() we can get some important statistics information, like min/max, mean, median, quartiles

```{r}
dim(df)
summary(df)
```

And we can plot a scatter of every columns combination, simply with pairs() or plot()

```{r}
plot(iris[,1:4],col=iris[,5],oma=c(1,4,6,12))
par(xpd=TRUE)
legend(0.85,0.6, as.vector(unique(iris$Species)),fill=c(1,2,3))
```

We can easily see some 


## Data Preprocessing

```{r}
df$Sepal.Length = scale( df$Sepal.Length )
df$Sepal.Width = scale( df$Sepal.Width )
df$Petal.Length = scale( df$Petal.Length )
df$Petal.Width = scale( df$Petal.Width )
summary(df)
```

We can split the data into test and train, using sample(list, n) to get n index number inside list. And then using the setdiff to get inside the list, numbers diferente from the train

```{r}
set.seed(420)
train = sample(1:150, 0.5*150) 
test <- setdiff(1:150,train)

df_train = iris[train,]

df_test <- iris[test,]
```

## Support Vector Machine

Using the library e1071, we can create an svm model

```{r}
svc = svm(Species ~ ., data = df_train)
summary(svc)
pred = predict(svc, df_test)
confusionMatrix(pred, df_test$Species)[3]$overall[1]
```

The predict function uses an model object, and then we can get the accuracy and an table to see how our model goes with the confusionMatrix

```{r}
pred = predict(svc, df_test)
confusionMatrix(pred, df_test$Species)[3]$overall[1]
```


L

